---
title: "Basic Operators"
---

## Arithmetic With Objects

All of the familiar arithmetic operators you practiced in the last tutorial (`+`, `-`, `*`, `/`) also work on named R
objects. When we do arithmetic on named objects, R substitutes the values stored inside those objects into the
expression before evaluating it.

```{r}
## First, define a few objects to use below.
x <- 2.5
y <- 7
z <- 5
```

**Do:** run the code above, then run the code below.

```{webr-r}
#| context: "setup"

x <- 2.5
y <- 7
z <- 5
```

```{webr-r}
y + x
y - x
y * x
y / x
```

**See:** each line prints a number, computed by substituting the stored values of `y` (`7`) and `x` (`2.5`) into the
expression.

::: practice

## Practice: Predict, Then Run

Given the objects defined above (`x = 2.5`, `y = 7`, `z = 5`), predict the value of each expression *before* running
it.

```{r, eval = FALSE}
z + y
z * x
z - y
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
z + y
z * x
z - y
```

## {{< var tab_title.solution >}}

```{r}
z + y
z * x
z - y
```

Notice that R never "sees" the letters `x`, `y`, `z` as text. Instead, it looks up whatever value is currently stored in each object and calculates with *that*. If you'd assigned different values to `x`, `y`, or `z`, these same three lines of code would produce entirely different answers without you changing a single character of the arithmetic itself. This is one of the most useful properties of programming: the same logic can be reused with different data.

::::

:::

## Why Bother With Objects at All?

You might wonder why we don't just write `7 + 2.5` directly instead of `y + x`. For a single calculation, writing the numbers directly is perfectly fine. But once your calculations get longer, or once you want to reuse a value multiple times, giving that value a name (an object) makes your code far easier to read, reuse, and fix.

We'll return to objects properly, including how to *create* them, a little later in this tutorial. For now, the important idea is just that R objects behave like named "slots" that hold a value, and arithmetic on those objects uses whatever value is stored inside.
