---
title: "Comments"
---

::: {.column-margin}

:::: reading

[R4DS 2.2: Comments](https://r4ds.hadley.nz/workflow-basics.html#comments)

::::

:::

Like any sensible programming language, R allows us to include comments in our code. The comment character in R is `#`.
Each line preceded by at least one `#` symbol will be interpreted as a comment. Comments are not evaluated when you run
your code. R simply ignores anything after the `#` character.

```{r}
# This is a comment, not code
1 + 1

2^2 # first three characters on this line are code; the rest is a comment
```

**Do:** Run the following code to generate two integer vectors and print the results.

```{webr-r}
1:10
2:8
```

**Predict:** What will the following lines print?

```{webr-r}
# 1:10
2:8
```

::: explanation

In the code above, the expression `1:10` is "commented out", so R doesn't evaluate that line at all. Consequently,
running the code only prints the result of the second expression, `2:8`.

:::

There are no block comments in R. So, you need to add a `#` character before every line when you want to create
multi-line comments.

```{r}
# This is a really long-winded explanation of exactly why the following is
# such a very super-duper awesome way of doing this thing that I'm trying to
# do here, because I'm a special little teapot, and I've got the very best
# ideas, and you really, really need to know about them.
x <- 1
y <- 1
z <- x + y
z
```

## Why Use Comments?

Comments have two primary uses.

**Use 1:** Leaving short, explanatory notes to yourself (or to anyone else reading your code) about *why* a particular
line does what it does.

```{r}
# Add 2 and 2
2 + 2
```

**Use 2:** Temporarily disabling parts of your code that you don't want to completely remove (e.g., for debugging).

For example, in the following code, I'm trying to accomplish a few things.

1. Generate a $25 \times 4$ matrix of standard normal deviates
1. Store this matrix as an object named `x`.
1. Compute the eigenvalue decomposition of `x`.
1. Calculate the covariance matrix of `x`.

**Do:** Run the following code, and see what happens.

```{webr-r}
x <- matrix(rnorm(100), ncol = 4)

xEigen <- eigen(x)

cor(x)
```

Oops...something breaks when trying to compute the eigenvalue decomposition. We'll assume I need that result, so I don't
want to simply remove the broken line. If I'm not willing or able to fix the issue immediately, I can comment out the
offending command and proceed with my work. Try it for yourself.

**Do:** In the interactive editor above, try commenting out the line `xEigen <- eigen(x)` and rerunning the code. 

**See:** Hopefully, the code now runs. As an added benefit, the commented line acts as a marker of the bug I need to fix
in the future.

::: knowledge-check

**Predict:** What will the following code print?

```{r, eval = FALSE}
5 + 5
# 5 - 5
5 * 5
```

Use the interactive editor to check your predictions.

```{webr-r}
```

:::
