---
title: "Comments"
---

::: {.column-margin}

:::: reading

[R4DS 2.2: Comments](https://r4ds.hadley.nz/workflow-basics.html#comments)

::::

:::

This is a short one. The comment character in R is `#`. Each commented line must be preceded by a `#` symbol. There are no block comments in R. Comments are not evaluated when you run your code. R simply ignores anything after a `#` on a given line.

**Do:** run the following R code to generate two integer vectors and print the results.

```{webr-r}
1:10
2:8
```

**Predict:** what do you expect to happen if you put a `#` in front of the first line below, then run the code?

```{webr-r}
#1:10
2:8
```

**Explain:** in the code above, the expression `1:10` is "commented out", so R doesn't evaluate that line at all. Instead, it's treated as plain text, not code. So, running the code only prints the second vector, `2:8`.

Comments are mainly useful for leaving short notes to yourself (or to anyone else reading your code) about *why* a particular line does what it does. You've already seen several examples of this throughout this module, e.g.:

```{r}
## Add 2 and 2
2 + 2
```

## Quick Check

What will the following code print? Predict first, then check your answer in the editor.

```{r, eval = FALSE}
5 + 5
# 5 - 5
5 * 5
```

```{webr-r}
```
