---
title: "Environments"
---

::: going-further

The information on this page is largely optional. It's good to have some awareness of R's global environment and how R
stores the objects you create, but you can safely progress to the next module without mastering these concepts.

:::

::: {.column-margin}

:::: reading-advanced

[Advanced R: Environments](https://adv-r.hadley.nz/environments.html)

::::

:::

Every R command is executed in a particular *environment*. Conceptually, you can think of R environments as collections
of the objects that R currently has stored in working memory. For most day-to-day R usage, you only need to consider one
of R's several types of environment: the *Global Environment*.

## The Global Environment

The global environment is the highest-level environment that R maintains. This environment contains all of the objects
you define in your R session. Whenever you create a new object (as you've been doing throughout this module), that
object lives in the global environment.

::: {.callout-note}

Since the global environment is the one you'll interact with most frequently, its often treated as the default
environment when economizing language (i.e., being lazy) or picking software defaults.

- In these tutorials, whenever we refer to "the environment", "your current environment", "your R session", or some
variation of those terms, we're talking about the global environment.
- The *Environment* tab in RStudio shows the contents of the global environment, by default.
- The interactions described below all interface with the global environment, by default.

:::

## Interacting with the (Global) Environment

We can use the `ls()` function to list the contents of the current environment.

```{r}
ls()
```

In this case, the environment is empty because we haven't yet defined any objects.

**Do:** Run the following code to add some objects to the current environment.

```{webr-r}
x <- 7
y <- 8
z <- 9
name <- "bob"
age <- 38
```

**Do:** Use the interactive editor below to run the `ls()` function. What do you see?

```{webr-r}
ls()
```

::: explanation

You should now see a listing of all the objects you just created. Remember: whenever we create a new object, we're
storing that object in the global environment.

:::

The `rm()` function will remove an object from the environment.

**Do:** Run the following code.

```{webr-r}
rm(x)
ls()
```

**See:** What effect did `rm(x)` have?

::: explanation

The listing returned by `ls()` no longer shows `x`, because you removed `x` from the environment by calling `rm(x)`.

:::

::: practice

1. Use the `ls()` function to view the contents of the environment.
1. Use the `rm()` function to remove `age` from the environment.
1. Use `ls()` to check your work.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
#| context: "setup"

y <- 8
z <- 9
name <- "bob"
age <- 38
```

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
#| include: FALSE

y <- 8
z <- 9
name <- "bob"
age <- 38
```

```{r}
ls()
rm(age)
ls()
```

::::

:::

::: {.callout-tip}

## Ready to Continue?

You're ready to move on if you can:

- Run a simple calculation in R. [Relevant Page](/tutorials/basic_commands/fundamentals.qmd)
- Create new R objects and use those objects for simple calculations. [Relevant
Page](/tutorials/basic_commands/objects.qmd)
- Write an expression that tests a logical condition. [Relevant Page](/tutorials/basic_commands/logical_comparisons.qmd)
- Explain what a comment does and write one yourself. [Relevant Page](/tutorials/basic_commands/comments.qmd)

If any of these concepts still feel shaky, that's completely fine. Revisit the relevant page before proceeding.

Take the [Knowledge Quiz](/quizzes/basic_commands/index.qmd) to evaluate your mastery of this module's learning goals.

:::
