---
title: "Environments"
---

::: {.callout-note}

## Optional Topic

This page is a bonus topic. You don't need to master environments to move on to the next tutorial on [R Functions](/tutorials/functions/index.qmd).

:::

## What is an "Environment"

::: {.column-margin}

:::: reading-advanced

[Advanced R: Environments](https://adv-r.hadley.nz/environments.html)

::::

:::

Every R command is executed in a particular *environment*. Conceptually, you can think of R environments as collections of the objects that R currently has stored in working memory. At this point, you only need to consider one of R's several
types of environment: the *Global Environment*.

### Global Environment

The global environment is the highest-level environment that R maintains. This environment contains all of the objects you define in your R session. The interactions described below all interface with the global environment. The *Environment* tab in RStudio shows the contents of the global environment, by default.

## Interacting with the Environment

We can use the `ls()` function to list the contents of the current environment.

```{r}
## The environment is empty because we haven't defined any objects
ls()

## Create some objects
x <- 7
y <- 8
z <- 9
name <- "bob"
age <- 38

## View the updated contents of the environment
ls()
```

The `rm()` function will remove an object from the environment. The following code will remove `x` from the environment.

```{r}
rm(x) # Remove 'x' from the environment
ls()  # Check the results
```

::: practice

## Practice

1. Use the `ls()` function to view the contents of the environment.
1. Use the `rm()` function to remove `age` from the environment.
1. Use `ls()` to check your work.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
#| context: "setup"

y <- 8
z <- 9
name <- "bob"
age <- 38
```

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
ls()
rm(age)
ls()
```
::::

:::

::: {.callout-tip}

## Ready to Continue?

You're ready to move on to [R Functions](/tutorials/functions/index.qmd) if you can:

- Run a calculation in R.
- Interpret the printed output of an R command.
- Write a simple R expression using at least one arithmetic or logical operator.
- Explain what a comment does, and write one yourself.

If any of these still feel shaky, that's completely fine, revisit the relevant page above before continuing, or try the [Knowledge Quiz](/quizzes/basic_commands/index.qmd) to see exactly where to focus.

:::
