---
title: "Additional Mathematical Operators"
---

Now that you're comfortable with the four basic arithmetic operators, we can extend your toolkit with a few additional,
commonly encountered mathematical operators. You don't need to memorize all of these today, treat this page as a
reference to which you can return.

```{r}
# Define three objects to use in the following examples
x <- 2.5
y <- 7
z <- 5
```

```{webr-r}
#| context: "setup"

x <- 2.5
y <- 7
z <- 5
```

## Powers & Roots

We use the caret character, `^`, to specify exponents. For example, the following code will square and cube the value of
`y`.

```{r}
y^2
y^3
```

The `sqrt()` function returns the square root of its argument (i.e., the value we specify inside the parentheses). The
following code calculates the square root of `y`.

::: {.column-margin}

:::: {.callout-note}

Don't worry about exactly what we mean by a *function*, just yet. You'll learn all about those in the [**R
Functions**](https://openstatprog.com/functions/index.html) module.

::::

:::

```{r}
sqrt(y)
```

For other roots, we can use fractional exponents.

```{r}
# Cute root of y
y^(1/3)

# Quartic root of y
y^(1/4) # quartic root of y
```

## Other Common Operators

R also includes many special functions for the most common mathematical operations.

```{r}
# Natural logarithm of 'y'
log(y)

# Base-10 logarithm of 'y'
log10(y)

# Base-2 logarithm of 'y'
log2(y)

# Exponentiate 'x'
exp(x)

# Modulo: Remainder after dividing 'y' by 'x'
y %% x

foo <- -3.14159

# Round 'foo' to 3 decimal places
round(foo, 3)

# Round 'foo' down to the nearest whole number
floor(foo)

# Round 'foo' up to the nearest whole number
ceiling(foo)

# Absolute value of 'foo'
abs(foo)
```

::: {.column-margin}

:::: {.callout-note}

Note that `log(y)` calculates the *natural logarithm* of `y`, $\ln(y)$. If you want the "ordinary" base-10 log, you need
to use the `log10()` function.

::::

:::

::: practice

## Practice

1. Create an object called `age` that takes the value of your age in whole years.
1. Use the `age` object you created above to create a second object called `weeks` that takes the value of your age
in whole weeks.
   - Assume 52 weeks in each year.
   - Disregard partial years (i.e., assume every year counted in `age` contains 52 whole weeks).
1. Print the value of `weeks`.
1. Use the modulo operator, `%%`, to check whether `weeks` is evenly divisible by 4.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

For step 4, remember that `a %% b` returns the remainder after dividing `a` by `b`. A value is evenly divisible by `b`
whenever that remainder is `0`.

## {{< var tab_title.solution >}}

At time-of-writing, I'm 40 years old. So, these would be my `age` and `weeks` objects.

```{r}
age   <- 40
weeks <- 40 * 52
weeks
```

To see if my `weeks` value is evenly divisible by 4, I'll apply the modulo operator.

```{r}
weeks %% 4
```

```{r}
#| include: FALSE

result <- ifelse(weeks %% 4 == 0, "is", "is not")
```

Since the remainder after dividing my `weeks` value by 4 `r result` equal to 0, we conclude that my `weeks` value
`r result` evenly divisible by 4.

::::

:::

::: knowledge-check

Which function would you use to find the remainder after division?

- `round()`
- `%%`
- `abs()`

Which function would you use to always round a number *up*?

- `floor()`
- `ceiling()`
- `round()`

Which function would you use to remove the negative sign from a number, without changing its magnitude?

- `sqrt()`
- `abs()`
- `exp()`

Try each candidate function in the editor to verify your answer.

```{webr-r}
```

:::
