---
title: "Extended Operators"
---

Now that you're comfortable with the four basic arithmetic operators and how R decides which operation to do first, we can extend your toolkit with a few more operators and functions that come up often in real analyses. You don't need to memorize all of these today, treat this page as a reference you can come back to.

```{r, echo = FALSE, include = FALSE}
## Define these objects so the static (knitr) chunks below can use them.
x <- 2.5
y <- 7
z <- 5
```

```{webr-r}
#| context: "setup"

x <- 2.5
y <- 7
z <- 5
```

## Powers & Roots

We use the caret character, `^`, to specify exponents. The following code will square and cube the value of `y`.

```{r}
y^2
y^3
```

**Do:** run the code, then predict the output of the next chunk before revealing it.

```{webr-r}
y^2
y^3
```

The `sqrt()` function returns the square root of its argument (i.e., the value we specify inside the parentheses). The
following code calculates the square root of `y`.

```{r}
sqrt(y)
```

For other roots, we can use fractional exponents.

```{r}
y^(1/3) # cube root of y
y^(1/4) # quartic root of y
```

## Other Common Operators

R also includes many special functions for the most common mathematical operations. Rather than trying to memorize
every one of these right now, skim the code and its comments once, then use it as a reference whenever you need one of
these operations later.

```{r}
log(y)   # natural logarithm of y
log10(y) # log base 10 of y
log2(y)  # log base 2 of y

exp(x) # exponentiate x

y %% x # modulo: remainder after dividing y by x

foo <- -3.14159

round(foo, 3) # round 'foo' to 3 decimal places
floor(foo)    # round 'foo' down to the nearest whole number
ceiling(foo)  # round 'foo' up to the nearest whole number

abs(foo) # absolute value of 'foo'
```

::: {.column-margin}

:::: {.callout-note}

Note that `log(y)` calculates the *natural logarithm* of `y`, $\ln(y)$. If you want the "ordinary" base-10 log, you need
to use the `log10()` function.

::::

:::

::: practice

## Practice

1. Create an object called `age` that takes the value of your age in whole years.
1. Use the `age` object you created above to create a second object called `weeks` that takes the value of your age
in whole weeks.
   - Assume 52 weeks in each year.
   - Disregard partial years (i.e., assume every year counted in `age` contains 52 whole weeks).
1. Print the value of `weeks`.
1. Use the modulo operator, `%%`, to check whether `weeks` is evenly divisible by 4.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

For step 4, remember that `a %% b` returns the remainder after dividing `a` by `b`. A value is evenly divisible by `b`
whenever that remainder is `0`.

## {{< var tab_title.solution >}}

At time-of-writing, I'm 38 years old. So, these would be my `age` and `weeks` objects.

```{r}
age   <- 38
weeks <- 38 * 52
weeks

weeks %% 4
```
::::

:::

## Quick Check

Which function would you use to...

- ...find the remainder after division? (`round()` / `%%` / `abs()`)
- ...always round a number *up*? (`floor()` / `ceiling()` / `round()`)
- ...remove the negative sign from a number? (`sqrt()` / `abs()` / `exp()`)

Try each candidate function in the editor to verify your answer if you're unsure.

```{webr-r}
```
