---
title: "Arithmetic"
---

## R as a Calculator

::: {.column-margin}

:::: reading

[R4DS 2.1: Coding Basics](https://r4ds.hadley.nz/workflow-basics.html#coding-basics)

::::

:::

At its simplest, R is a very over-powered pocket calculator. Run the code below (click the small "Run Code" button, or
press `Ctrl`/`Cmd` + `Enter`) and look at the output.

```{webr-r}
2 + 2
```

Congratulations! Assuming you saw the answer `4`, you've successfully used R to sum 2 and 2.

::: practice

Using the interactive editor below, add your birth day to your birth month.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

Since I was born on 26 August, my solution would be the following.

```{r}
26 + 8
```

::::

:::

In R, the four basic arithmetic operators all behave the way you'd expect.

```{r}
# Add 2 and 2
2 + 2

# Subtract 6 from 14
14 - 6

# Multiply 3 and 4
3 * 4

# Divide 225 by 15
225 / 15
```

::: practice

Before running any code, **predict** what value each of the following expressions will produce, *then* use the
interactive editor to run the code and check your predictions.

```{r, eval = FALSE}
10 - 3
4 * 4
100 / 4
```

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
10 - 3
4 * 4
100 / 4
```

If your predictions matched, great! You've already got the core idea. If not, that's completely fine. Notice *which*
line surprised you, make sure you correctly identified the arithmetic operator involved, and try to work out where your
prediction went wrong.

::::

:::

Now let's use these same tools to solve a slightly more complex problem.

::: practice

1. Calculate the number of whole days left in the current month.
1. Calculate the number of weeks left in the month.
   - Account for partial weeks with a decimal-valued answer.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

At time-of-writing, it's 21 August. Since August has 31 days, we get 31 - 21 = 10 days remaining in the month.

```{r}
31 - 21
```

To find the number of remaining weeks, we simply divide the answer from above by 7.

```{r}
10 / 7
```

::::

:::

## Order of Operations

When parsing your commands, R will, mostly, scan each line of code from left to right and apply each mathematical
operation according to the usual PEMDAS ordering (**P**arentheses, **E**xponents, **M**ultiplication/**D**ivision,
**A**ddition/**S**ubtraction).

**Do:** Run the following two lines of code and compare their results.

```{webr-r}
3 + 4 / 2
(3 + 4) / 2
```

**See:** Notice that the two lines produce different answers, even though they use the same three numbers and the same
three operators. Why?

- In the first line, R first divides `4` by `2`, and *then* adds `3` to the result.
- In the second line, the parentheses force R to first add `3` to `4`, and only *then* divide the result by `2`.

Parentheses are your most reliable tool for controlling order of operations. When in doubt, add parentheses. They never
hurt, and they make your intent explicit both to R and to anyone (including future-you) reading the code.

::: practice

Why do the following two expressions produce different answers?

:::: {.panel-tabset}

## {{< var tab_title.question >}}

```{r}
(3 + 4 - 2) * (10 - 7)
3 + 4 - 2 * 10 - 7
```

## {{< var tab_title.solution >}}

```{r}
(3 + 4 - 2) * (10 - 7)
```

In the first case, we've used parentheses to group the two blocks of addition/subtraction $(3 + 4 - 2 = 5)$ and $(10 - 7
= 3)$. We then multiply those results $(5 \times 3 = 15)$.

```{r}
3 + 4 - 2 * 10 - 7
```

In the second case, we rely on R's rules for operator precedence, so the multiplication $(2 \times 10 = 20)$ is done
first. Then we do the addition/subtraction $(3 + 4 - 20 - 7 = -20)$.

::::

:::

Now that you're aware of R's rules for operator precedence, let's revisit the birth day/month problem from above.

::: practice

Use a single line of code to complete the following calculations.

1. Calculate the number of whole days left in the current month.
1. Calculate the number of weeks left in the month.
   - Account for partial weeks with a decimal-valued answer.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

The calculation is the same as last time, but now we can combine the remaining days difference and the remaining weeks
division into one expression.

```{r}
(31 - 21) / 7
```

Note the use of parentheses to ensure that we compute the number of remaining days *before* doing any division.

::::

:::

::: knowledge-check

Would the following two expressions produce the same result?

```{r}
#| eval: FALSE

(3 * 4) - (10 / 2)
3 * 4 - 10 / 2
```

Use the interactive editor below to check your prediction.

```{webr-r}
```

:::
