---
title: "Logical Comparisons"
---

We can test many flavors of logical conditions in R. Logical tests return a logical value as the result. A logical value takes the value of `TRUE` whenever the tested condition is satisfied and `FALSE` when the tested condition is not satisfied.

```{webr-r}
#| context: "setup"

y <- 5
x <- 7
w <- 5
```

## Equality

The simplest logical test is an equality check. To check if two objects are equal, we use the 'equality operator', `==`.

```{r}
## Define some objects to play with
y <- 5
x <- 7
w <- 5
```

```{webr-r}
### Check equality
y == x
y == w
```

**Explain:** `y == x` is `FALSE` because `5` is not equal to `7`. `y == w` is `TRUE` because both objects store the value `5`. Notice `==` (two equals signs) is a *comparison*, while `<-` and a single `=` are used for *assignment*. Mixing these up is one of the most common beginner mistakes in R.

## Greater/Lesser

We can also check the usual greater-than/less-than conditions with `>`, `<`, `>=`, `<=`.

```{r}
y > x  # greater than
y >= x # greater than or equal to
y < x  # less than
y <= x # less than or equal to
```

::: practice

## Practice: Predict, Then Run

Given `y <- 5` and `w <- 5`, predict what value each of the following expressions will return.

```{r, eval = FALSE}
y > w
y >= w
y < w
y <= w
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
y > w
y >= w
y < w
y <= w
```

## {{< var tab_title.solution >}}

```{r}
y > w
y >= w
y < w
y <= w
```

Since `y` and `w` store the *same* value, `y > w` and `y < w` are both `FALSE` (neither is strictly greater or less than the other), while `y >= w` and `y <= w` are both `TRUE`, since "greater than or equal to" and "less than or equal to" are satisfied whenever the values are equal.
::::

:::

## Logical Negation

We can negate any logical condition by prepending the `!` character.

```{r}
y > x
!y > x

y == w
!y == w
```

Rather than negating an equality check, we will typically use the special "not-equal" operator, `!=`.

```{r}
y == w
y != w
```

## Combining Conditions

We can create more complex logical conditions with the AND and OR operators: `&` and `|`.

```{r}
y == w & y < x
y == w & y > x
y == w | y > x
```

::: practice

## Practice

Use a single line of code to generate a logical value (i.e., TRUE/FALSE) indicating if the value of the `weeks` object you created earlier is evenly divisible by 7.

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.hint >}}

- Use `x %% 7` to calculate the remainder after dividing `x` by `7`.
- If you're getting unexpected results, consider using parentheses to ensure the intended order of operations.
- In R, the modulo operator, `%%`, takes precedence over arithmetic operations like multiplication.

## {{< var tab_title.solution >}}

```{r, echo = FALSE, include = FALSE}
weeks <- 38 * 52
```

```{r}
(weeks %% 7) == 0
```

If the remainder of dividing `weeks` by 7 is zero, we know that the value of `weeks` is evenly divisible by 7.
::::

:::

## Quick Check

- What's the difference between `=` and `==` in R?
- If `a <- 3` and `b <- 4`, what does `a != b` return?
- Write (but don't necessarily run) an expression that checks whether `a` is greater than `1` *and* less than `10`.

```{webr-r}
```
