---
title: "Logical Comparisons"
---

```{r}
#| include: FALSE

y <- 5
x <- 7
w <- 5
```

Almost all the commands you've seen so far have been some sort of mathematical operation (e.g., adding, dividing,
computing the natural logarithm), but R can do much more for us. *Logical comparisons* are one particularly useful class
of operations, in which we compare objects to test logical conditions.

We can test many flavors of logical conditions in R. Logical tests return a logical value as the result. A logical value
is `TRUE` whenever the tested condition is satisfied and `FALSE` when the tested condition is not satisfied.

::: {.column-margin}

:::: going-further

The "logical values" we describe here are actually length-one logical vectors. You'll learn all about atomic vectors in
the [**Vectors & Matrices**](https://openstatprog.com/vectors-matrices/index.html) module.

::::

:::

**Do:** Run the following code to create three objects that we can use below.

```{webr-r}
#| autorun: true

y <- 5
x <- 7
w <- 5
```

## Equality

The simplest logical test is an equality check. To check if two objects are equal, we use the equality test operator,
`==`.

```{r}
1 == 1
"a" == "a"
"foo" == "bar"
```

::: common-mistakes

Notice that the equality test operator, `==`, is two equals signs. Remember that a single equals sign is one of the
possible assignment operators. You cannot use a single equals sign to do a logical comparison.

```{r}
#| error: TRUE

2 = 3
```

Mixing up `==` and `=` is one of the most common beginner mistakes (and experienced user typos) in R.

:::

**Predict:** What will the following lines print?

```{webr-r}
y == x
y == w
```

::: explanation

- `y == x` is `r y == x` because `r y` *`r ifelse(y == x, "is", "is not")`* equal to `r x`.
- `y == w` is `r y == w` because `r y` *`r ifelse(y == w, "is", "is not")`* equal to `r w`.

:::

## Greater-Than & Less-Than

We can also check the usual greater-than/less-than conditions with `>`, `<`, `>=`, `<=`.

```{r}
# Is y greater than x?
y > x

# Is y greater than or equal to x?
y >= x

# Is y less than x?
y < x

# Is y less than or equal to x?
y <= x
```

::: practice

Given `y <- 5` and `w <- 5`, **predict** what value each of the following expressions will return.

```{r, eval = FALSE}
y > w
y >= w
y < w
y <= w
```

Use the interactive editor to check your predictions.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
y > w
y >= w
y < w
y <= w
```

`y` and `w` store the *same* value, so:

- `y > w` and `y < w` are both `FALSE`, since neither is strictly greater or less than the other.
- `y >= w` and `y <= w` are both `TRUE`, since "greater than or equal to" and "less than or equal to" are satisfied
whenever the values are equal.

::::

:::

## Logical Negation

We can negate any logical condition (i.e., change `TRUE` to `FALSE` and visa-versa) by prepending the `!` character.

```{r}
!TRUE
!FALSE

y > x
!y > x
```

**Predict:** What will the following two lines print?

```{webr-r}
y == w
!y == w
```

::: explanation

- `y == w` is `r y == w` because `r y` *`r ifelse(y == w, "is", "is not")`* equal to `r w`.
- `!y == w` is `r !y == w` because we've negated the result of the equality test.

:::

Rather than negating an equality check, we will typically use the special "not-equal" operator, `!=`.

```{r}
y == w
y != w
```

## Combining Conditions

We can create more complex logical tests by combining multiple conditions with the AND and OR operators: `&` and `|`.

Logical AND tests if all conditions are satisfied.

- $A$ AND $B$ is true only when both Condition A and Condition B are satisfied.
- $A$ AND $B$ is when Condition A, Condition B, or both are *not* satisfied.

```{r}
TRUE & TRUE
TRUE & FALSE
FALSE & FALSE
```

Logical OR tests if at least one condition is satisfied.

- $A$ OR $B$ is true when Condition A, Condition B, or both are satisfied.
- $A$ OR $B$ is false only when both Condition A and Condition B are *not* satisfied.

```{r}
TRUE | TRUE
TRUE | FALSE
FALSE | FALSE
```

To run multi-condition tests, we simply build up the desired logic by chaining together each constituent condition.

```{r}
# Is y equal to w AND is y less than x?
y == w & y < x

# Is y equal to w AND is y greater than x?
y == w & y > x

# Is y equal to w OR is y greater than x?
y == w | y > x
```

**Predict:** What will the following lines print?

```{webr-r}
y == y & y != y
w > w | w < w
x > x | x <= x
x >= x & x <= x
```

::: explanation

1. `FALSE` because `y` is not both equal to itself and not equal to itself.
1. `FALSE` because `w` is neither greater than itself nor less than itself.
1. `TRUE` because `x` is not greater than itself, but `x` is less than or equal to itself.
1. `TRUE` because `x` is both greater than or equal to itself and `x` is less than or equal to itself.

:::

::: practice

Use a single line of code to generate a logical answer (i.e., `TRUE`/`FALSE`) to the following question.

- Is the sum of the current day and the current month strictly greater than the two-digit year?

Use numeric representations of the day, month, and year.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

Let's assume the current date is 25-08-2026. Then, today's solution would be the following.

```{r}
(25 + 8) > 26
```

In R, most mathematical operators take precedence over logical tests. So, the parentheses in the solution are not
strictly necessary.

```{r}
25 + 8 > 26
```

That being said, the parentheses are still a good idea. In this case, the extra parentheses clarify exactly what we're
testing, and less ambiguity is almost always a good thing.

::::

:::

::: knowledge-check

- What's the difference between `=` and `==` in R?
- If `a <- 3` and `b <- 4`, what does `a != b` return?
- Write an expression that checks whether `a` is greater than `1` *and* less than `10`.

```{webr-r}
```

:::
