---
title: "Objects (Assignment)"
---

## Why Do We Need Objects?

::: {.column-margin}

:::: reading

[R4DS 2.3: Object Names](https://r4ds.hadley.nz/workflow-basics.html#sec-whats-in-a-name)

::::

:::

::: {.column-margin}

:::: reading-advanced

[Advanced R 2.2: Binding Basics](https://adv-r.hadley.nz/names-values.html#binding-basics)

::::

:::

So far, every value we've used has disappeared as soon as R printed it, we haven't *saved* anything. To do anything useful with R, we need a way to store a value so we can use it again later. We do this by creating **objects** that hold data.

## Creating Objects

We 'assign' values to objects via the 'assignment' operator, `<-`.

The following code assigns the values `7`, `2.5`, and `"foo"` to the objects `x`, `y`, and `z`, respectively.

```{r}
x <- 7
y <- 2.5
z <- "foo"
```

**Do:** run the code above.

**See:** notice the lack of printed output. These three commands create three new objects in your environment (`x`, `y`, `z`) that store the assigned data values (`7`, `2.5`, `"foo"`). However, we haven't yet asked R to do anything with those stored values, so we don't see any printed output.

**Explain:** assignment is a *silent* action: it changes what R remembers, but it doesn't print anything by itself. This trips up almost everyone the first time, so it's worth internalizing now: *assigning* a value and *displaying* a value are two separate steps.

## Viewing an Object's Value

To view the contents of an object, we can evaluate the object's name without assignment.

```{webr-r}
#| context: "setup"

x <- 7
y <- 2.5
```

Run the following code to print the values saved as `x` and `y`.

```{webr-r}
x
y
```

::: practice

## Practice: Predict, Then Run

Given `x <- 7` and `y <- 2.5` (already run above), predict what each line below will print before running it.

```{r, eval = FALSE}
x + y
x
x - 1
x
```

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}
x + y
x
x - 1
x
```

## {{< var tab_title.solution >}}

```{r}
x + y
x
x - 1
x
```

Notice that `x` still equals `7` after the last line, even though we just computed `x - 1`. Doing arithmetic *with* an object does not change the value stored *in* that object, R only updates an object's value when you explicitly re-assign it with `<-`.
::::

:::

## Copies, Not Renames

We can also create new objects by assigning them the values of existing objects.

The following code creates a new object, `w`, that takes the value of the existing object `z`. When we print the value
of `w`, you can see that it has the same value as `z`.

```{r}
w <- z
w
z
```

When we create `w` above, we are not replacing `z` with `w`: we're making a copy of `z` and calling that copy `w`. So, both `w` and `z` are still available. In almost all cases, R will copy objects during assignment. This behavior is good to keep in mind when you're working with larger datasets: you can quickly flood your memory with unnecessary copies of your data, if you're not careful.

## Quick Check

- After running `w <- z`, does changing `w` also change `z`? Why or why not?
- What is printed when you run `x <- 7` by itself? What about when you run just `x` on the next line?

Test both questions in the editor if you're not sure.

```{webr-r}
```
