---
title: "R Objects & Assignment"
---


On the last page, every value we used disappeared as soon as R printed it, we didn't *store* any data for later. To do
anything useful with R, we need a way to store a value so we can use it again in future commands. We do this by creating
**objects** that hold the pieces of data we want to work with.

::: {.column-margin}

:::: reading

[R4DS 2.3: Object Names](https://r4ds.hadley.nz/workflow-basics.html#sec-whats-in-a-name)

::::

:::

::: {.column-margin}

:::: reading-advanced

[Advanced R 2.2: Binding Basics](https://adv-r.hadley.nz/names-values.html#binding-basics)

::::

:::

## Creating Objects

We create objects by 'assigning' values to names via the 'assignment' operator, `<-`. The following code assigns the
values `7`, `2.5`, and `"foo"` to the objects `x`, `y`, and `z`, respectively.

**Do:** Run the following code.

```{webr-r}
#| autorun: TRUE

x <- 7
y <- 2.5
z <- "foo"
```

**See:** Notice the lack of printed output. These three commands create three new objects in your environment (`x`, `y`,
`z`) that store the assigned data values (`7`, `2.5`, `"foo"`). However, we haven't yet asked R to do anything with
those stored values, so we don't see any printed output.

Assignment is a *silent* operation: it creates a new object in the current R session but doesn't print anything. This
behavior can be confusing, so it's worth taking a moment to internalize the distinction. *Assigning* a value and
*printing* a value are two separate steps.

## Printing an Object's Value

To view the contents of an object, we can evaluate the object's name without assignment. Doing so will print the
contents of the object in the R console.

**Do:** Run the following code to print the values saved as `x` and `y`.

```{webr-r}
x
y
```

::: {.callout-note}

Technically, `=` and `->` can also be used for assignment, but doing so is strongly discouraged by almost all style
guides.

```{r}
# Don't do this
a = "alice"
a

# Don't do this either
"bob" -> b
b
```

Unless you have a very good reason to do otherwise, use only `<-` for assignment.

:::

## Arithmetic with Objects

All of the familiar arithmetic operators you practiced on the previous page (`+`, `-`, `*`, `/`) also work on named R
objects. When we do arithmetic on named objects, R substitutes the values stored inside those objects into the
expression before evaluating it.

```{r}
# First, define a few objects to use below.
x <- 2.5
y <- 7
z <- 5
```

**Do:** Run the following code to create the objects `x`, `y`, and `z`.

```{webr-r}
x <- 2.5
y <- 7
z <- 5
```

**See:** Notice that these assignment commands don't print anything.

**Do:** Run the following code and note what each line produces.

```{webr-r}
y + x
y - x
y * x
y / x
```

**See:** Each line prints a number, computed by substituting the stored values of `y` (`7`) and `x` (`2.5`) into the
expression.

::: practice

Given `x <- 7` and `y <- 2.5` (already run above), predict what each of the following lines will print before running
them.

```{r}
#| eval: FALSE

x
y
x + y
x - 1
x
```

Use the interactive editor to check your predictions.

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
x
y
x + y
x - 1
x
```

::::: {.callout-important}

Notice that `x` still has the value `7` when running the last line, even though we just computed `x - 1` on the line
above. Doing arithmetic *with* an object does not change the value stored *in* that object. R only updates an object's
value when you explicitly re-assign it with `<-`.

:::::

::::

:::

## Assigning Existing Objects

In real-world data analysis, we rarely create objects by directly assigning single numbers. Typically, we create objects
by manipulating existing objects and assigning the manipulated version to a new name. Fortunately, R is perfectly happy
to create new objects by assigning them the values of existing objects.

The following code creates a new object, `w`, that takes the value of the existing object `z`. When we print the value
of `w`, you can see that it has the same value as `z`.

```{r}
w <- z
w
z
```

::: {.callout-important}

When we create `w` in the code above, we are not replacing `z` with `w`: we're making a copy of `z` and naming that copy
`w`. So, both `w` and `z` are still available (otherwise we wouldn't be able to print the value of `z`). In almost all
cases, R will copy objects during assignment. This behavior is good to keep in mind as you progress to more complicated
projects. If you're not careful, your environment can quickly become cluttered with unused copies of old objects.

:::

::: practice

Given the objects defined above (`x <- 2.5`, `y <- 7`, `z <- 5`), predict the value of each expression *before* running
it.

```{r}
#| eval: FALSE

z + y
z * x
z - y
```

:::: {.panel-tabset}

## {{< var tab_title.editor >}}

```{webr-r}
```

## {{< var tab_title.solution >}}

```{r}
z + y
z * x
z - y
```

Notice that R does not interpret the letters `x`, `y`, `z` as text. Instead, it uses whatever value is currently stored
in each object when doing the calculations. If you assign different values to `x`, `y`, or `z`, these same three lines
would produce entirely different answers. Try it for yourself.

In the following code block, replace the `?` with a new value for `z`, and rerun the expressions.

```{webr-r}
z <- ?

z + y
z * x
z - y
```

Notice how each line now produces a different result, even though you didn't change anything about the expressions,
themselves. This paradigm is one of the most useful properties of programmatic data analysis. By changing the values
stored in the relevant objects (e.g., by loading a different dataset), you can reuse the same code to re-run the same
calculation on different data.

::::

:::

::: common-mistakes

Even though we name objects with words, R has different ways of handling named objects and character strings.

- We use quotation marks to denote a character string.
- We don't quote object names.

If we run `foo <- "bar"`, `foo` is an object that stores the character string `"bar"`. We need to be careful with what
we quote: `foo` and `"foo"` are completely different entities. Most quotation-based mistakes fall into one of two
classes.

**Mistake 1:** We forget the quotation marks when trying to define a character string.

```{r}
#| error: TRUE

# We want to assign the character string "bar" to the object foo
foo <- bar
```

Since we didn't quote the value on the right-hand-side of the assignment operator, R will try to find an object called
`bar` and assign its contents to `foo`. Since, `bar` doesn't exist, we get an error.

**Mistake 2:** We quote an objects name.

```{r}
#| error: TRUE

alice <- 42
bob <- 1984

# We want to add the values stored in the objects alice an bob
alice + "bob"
```

In this case, we should have written `alice + bob`. Both `alice` and `bob` exist as objects that store numeric values,
but `"bob"` is just a character string that has no relation to the object `bob`. So, R tries to add the value stored in
`alice` (i.e., `r alice`) to the string-literal `"bob"` and fails.

:::

## Why Bother?

You might wonder why we don't just write `7 + 2.5` directly instead of first assigning `7` to `y`, then assigning `2.5`
to `x`, and finally computing `y + x`. For a single calculation, writing the numbers directly is perfectly fine, but not
when your calculations get longer and more complex. In real-world projects, you will need to reuse certain data values
multiple times. So, storing those values as named objects makes your code far easier to read, reuse, and fix.

::: knowledge-check

- What is printed when you run `x <- 7` by itself? What about when you run just `x` on the next line?
- After running `w <- z`, does changing `w` also change `z`? Why or why not?

Use the interactive editor to check your understanding.

```{webr-r}
```

:::
