---
title: "Order of Operations"
---

## PEMDAS in R

When parsing your commands, R will, mostly, scan each line of code from left to right and apply each mathematical
operation according to the usual PEMDAS ordering (**P**arentheses, **E**xponents, **M**ultiplication/**D**ivision,
**A**ddition/**S**ubtraction).

```{r, echo = FALSE, include = FALSE}
## Define these objects so the static (knitr) chunks below can use them.
x <- 2.5
y <- 7
z <- 5
```

```{webr-r}
#| context: "setup"

x <- 2.5
y <- 7
z <- 5
```

**Do:** run both lines below and compare the results.

```{webr-r}
y + x / z
(y + x) / z
```

**See:** the two lines produce different answers, even though they use the same three numbers and the same three
operators.

**Explain:**

- In the first line, R first divides `x` by `z`, and *then* adds `y` to the result.
- In the second line, the parentheses force R to first add `x` to `y`, and only *then* divide the result by `z`.

Parentheses are your most reliable tool for controlling order of operations. When in doubt, add parentheses. They never hurt, and they make your intent explicit both to R and to anyone (including future-you) reading the code.

::: practice

## Practice

Why do the following two expressions produce different answers? Before running anything, write down what each line is asking R to compute, step by step.

:::: {.panel-tabset}
## {{< var tab_title.question >}}

```{r}
y^(1/2)
y^1/2
```

## {{< var tab_title.solution >}}

```{r}
y^(1/2)
sqrt(y)
```
In the first case, we're raising `y` to the power of $\frac{1}{2}$ — which is the same as taking a square root.

```{r}
y^1/2
y/2
```
In the second case, R applies the exponent (`^`) *before* the division, so it first raises `y` to the power of `1`
(which does nothing) and only then divides the result by `2`. The exponent operator has higher precedence than
division, so it's applied first even though it appears "attached" to the `1` rather than the whole expression.
::::

:::

## Quick Check

- If you wanted `y` raised to the power of `1/2`, which version is correct: `y^(1/2)` or `y^1/2`?
- What single change would make `y^1/2` compute the same thing as `y^(1/2)`?

Try both in the editor below to confirm your reasoning.

```{webr-r}
```
